---
title: "TN School Stats"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
Inputting school data from Tennessee State Performance Tests to see how does discipline and chronic absenteeism affect performance?
```{r setup}
#install.packages("DBI")
library(tidyverse)
library(DBI)
library(readr)
db <- dbConnect(RSQLite::SQLite(), dbname = "sql.sqlite")

setwd("~/Downloads/School/SQL")

#Inputting the tables of only school values and making a merge key for joining later 
tenn_18 <- read_csv("tenn_2018_achieve_copy.csv")  %>% 
  filter(!is.na(school_name)) %>% 
  unite(dis_school, district_number, school_name, sep = "_") 
dis <-read_csv("discipline_tn_copy.csv") %>% 
   filter(!is.na(school_name)) %>% 
    unite(dis_school, district_number, school_name, sep = "_") 
ca <-read_csv("chronic_absent_copy.csv") %>% 
   filter(!is.na(school_name)) %>% 
    unite(dis_school, district_number, school_name, sep = "_") 
```


#Writing the Tables to SQL database
```{r write data}
dbWriteTable(db,
    "tenn_18",
    tenn_18, # df being written
    overwrite = TRUE)
dbWriteTable(db, 
    "dis",
    dis, # df being written 
    overwrite = TRUE)
dbWriteTable(db, 
    "ca",
    ca, # df being written 
    overwrite = TRUE)
```

#Ensursing our tables were created in SQL
```{sql, connection=db}
SELECT * FROM sqlite_master;
```

#Previewing the tenn_18 table
```{sql, connection=db}
SELECT *
FROM "tenn_18"
LIMIT 10
```

#What is the count of distinct schools in the dataset?
```{sql, connection=db}
SELECT COUNT(DISTINCT dis_school) FROM "tenn_18";
```

#Joining the tables to measure discipline or chronic absenteeism 
```{sql, connection=db}
CREATE TABLE tenn_18_dis AS
SELECT * 
FROM tenn_18 
LEFT JOIN dis 
ON(tenn_18.dis_school = dis.dis_school);
```

#Previewing the tenn_18_dis table
```{sql, connection=db}
SELECT *
FROM "tenn_18_dis"
LIMIT 10
```

#Joining the tables to measure discipline or chronic absenteeism 
```{sql, connection=db}
CREATE TABLE tenn_18_ca AS
SELECT * 
FROM tenn_18 
LEFT JOIN ca
ON(tenn_18.dis_school = ca.dis_school);
```

#Previewing the tenn_18_ca table
```{sql, connection=db}
SELECT *
FROM "tenn_18_ca"
LIMIT 10
```

#What are the schools with the highest percent on mastered test scores?
```{sql, connection=db}
SELECT percent_on_mastered, school_name
FROM "tenn_18"
WHERE subgroup = "All Students"
AND school_name IS NOT NULL
GROUP BY school_name
ORDER BY percent_on_mastered DESC
```
#What are the schools with the lowest percent on mastered  test scores?
```{sql, connection=db}
SELECT percent_on_mastered, dis_school
FROM "tenn_18"
WHERE subgroup = "All Students"
AND percent_on_mastered IS NOT NULL
GROUP BY dis_school
ORDER BY percent_on_mastered ASC
```

#What the schools with the highest chronic absenteeism?
```{sql, connection=db}
SELECT percent_ca, dis_school
FROM "ca"
WHERE subgroup = "All Students"
AND percent_ca != '*'
GROUP BY dis_school
ORDER BY percent_ca DESC
```

#What the schools with the highest discipline percentage?
```{sql, connection=db}
SELECT percent, dis_school
FROM "dis"
WHERE subgroup = "All Students"
AND percent IS NOT NULL
GROUP BY dis_school
ORDER BY percent DESC
```

#Selecting the lowest performing school and subject
```{sql, connection=db}
SELECT MIN(percent_on_mastered), dis_school, overall_subject
FROM "tenn_18"
WHERE subgroup = "All Students"
```
#Selecting the highest performing school and subject
```{sql, connection=db}
SELECT MAX(percent_on_mastered), dis_school, overall_subject
FROM "tenn_18"
WHERE subgroup = "All Students"
```
#What is the chronic absenteeism at the lowest performing school?
```{sql, connection=db}
SELECT MIN(percent_on_mastered), dis_school, percent_ca
FROM "tenn_18_ca"
WHERE subgroup = "All Students"
```
##What is the discipline rate at the lowest performing school?
```{sql, connection=db}
SELECT MIN(percent_on_mastered), dis_school, percent
FROM "tenn_18_dis"
WHERE subgroup = "All Students"
```

#What is the chronic absenteeism at the highest performing school?
```{sql, connection=db}
SELECT MAX(percent_on_mastered), dis_school, percent_ca
FROM "tenn_18_ca"
WHERE subgroup = "All Students"
```
##What is the discipline rate at the highest performing school?
```{sql, connection=db}
SELECT MAX(percent_on_mastered), dis_school, percent
FROM "tenn_18_dis"
WHERE subgroup = "All Students"
```
- The lowest performing school had a high chronic absenteeism of 18.6 percent and .2 percent discipline rate, while the high performing school had a 3.5 percent of chronic absenteeism and a 0 percent rate of discipline.

#What is the average science score across all school?
```{sql, connection=db}
SELECT AVG(percent_on_mastered)
FROM "tenn_18"
WHERE subgroup = "All Students"
AND school_name IS NOT NULL
AND overall_subject = "Science"
```

#What is the average chronic absenteeism across all school?
```{sql, connection=db}
SELECT AVG(percent_ca)
FROM "ca"
WHERE subgroup = "All Students"
```

#Case when to find outliers of disciplinary measures
```{sql, connection=db}
SELECT 
percent_ca, 
dis_school,
CASE WHEN percent_ca > 12.23608	 THEN 'above average absenteeism'
            WHEN percent_ca < 12.23608	  THEN 'below average'
            WHEN percent_ca = '*' THEN 'na'
            ELSE "average absenteeism"
            END AS absenteeism
FROM ca;
```